Title

### 2012-08-29 inc/laser/laser.tex
removed some errant unicode

% page 8630
Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium’s advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load. 
Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.

@@

The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
%TODO change voice/check ->
To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
%TODO replace to "This output will then be use to ..." and link to something from previous
%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} – a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today’s ultrashort pulse laser systems.
%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance\cite{zawadzka_evanescent_2001}.
%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved\cite{zawadzka_evanescent_2001}.

@@

In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH\cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig. 4(b)) at an optimum mode-locked output power of $5.0$W.
This trend is consistent with the expected ‘soliton’ modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.

@@

Figure 5 shows the laser's output power as a function of diode current for the cavity configuration depicted in Fig. 3(a); a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (cw) output power of 1.8W.
Between diode currents (output powers) of 23 and 25A (1.8 and 2.3W), the resonator operates mainly in the cw mode, but will transition to ML operation upon being subjected to a perturbation.
The switching to ML operation is accompanied by a 250mW increase in output
%page 8637
power indicating a higher cavity Q for ML operation.
This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal’s optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
From Fig. 3(b), we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.

@@

The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster’s angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.
At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
After the collimating
%page 8639
mirror, a standard flat dichroic beamsplitter then separates the reflected $s$-polarized 523nm radiation from the transmitted $p$-polarized fundamental.

>>>>

% page 8630
Ytterbium-doped potassium gadolinium tungstate (\ce{Yb$:$KGd(WO4)2} or Yb:KGW) is becoming a widely used solid-state gain medium for ultrafast laser systems.
As with other Yb-doped laser crystals\cite{Brenier_new_criteria}, this is primarily due to the gain medium's advantageous absorption properties for direct laser diode pumping at 980nm, its broad emission bandwidth around 1040nm, and consequent small ($\sim$ 6\%) quantum defect that results in a relatively low thermal load. 
Mode-locked (ML), diode-pumped, Yb:KGW laser oscillators generating pulses with sub-picosecond durations at average output powers in excess of 1W have been demonstrated\cite{Brunner_diode_pumped,Courjaud_high_power,Major_femtosecond_2006,Holtom_mode_locked_2006} and are now available commercially\cite{website_amplitude,website_solar}.
Such lasers have used dichroic mirrors\cite{Brunner_diode_pumped,Major_femtosecond_2006,Paunescu_diode_2004,Major_extended_2006}, polarization-coupling\cite{Holtom_mode_locked_2006}, or more complex optical geometries (e.g., thin disk lasers\cite{website_amplitude,Brunner_pulses_2002}) to pump the laser crystal in an efficient manner.

@@

The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
%TODO change voice/check ->
To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.
%TODO replace to "This output will then be use to ..." and link to something from previous
%In our case, the visible femtosecond pulses will be used as the driving laser radiation source for an ultrafast electron microscope \cite{king_ultrafast_2005} --- a technology that seeks to combine the sub-nanometer spatial resolution of electron microscopy with the sub-picosecond temporal resolution afforded by today's ultrashort pulse laser systems.
%Critical to this technology is likely to be the development of a femtosecond photo-electron gun with a low spatial emittance\cite{zawadzka_evanescent_2001}.
%The coherent control of photoemission through visible laser-driven plasmon excitations on large-area (a few mm2), gold or silver, nano-patterned photocathodes may provide a mechanism by which this can be achieved\cite{zawadzka_evanescent_2001}.

@@

In the cavity configuration depicted in Fig. 3(a), self-starting mode-locked (ML) operation was observed for output powers greater than $\sim$2W and $z \approx 42$cm.
The maximum (or optimum) mode-locked output power obtained without the emergence of DC spectral components was 3.5W, with a sech$^2$ pulse duration of 250fs (Fig. 4(a)).
The insertion of two plane dispersion compensation mirrors each with a GVD of $-700\text{fs}^2$ (Layertec GmbH\cite{website_layertec}) into the short arm of the cavity generated higher mode-locked output powers with longer pulse durations.
For example, with two reflections off each mirror (an additional $-5600\text{fs}^2$ of round-trip GVD), the pulse duration increased to 347fs (Fig. 4(b)) at an optimum mode-locked output power of $5.0$W.
This trend is consistent with the expected `soliton' modelocking regime; that is, the SBR only initiates the mode-locking mechanism, which is quickly dominated by self-phase modulation in the Yb:KGW gain medium balanced by the net negative GVD in the cavity\cite{Brunner_diode_pumped}.

@@

Figure 5 shows the laser's output power as a function of diode current for the cavity configuration depicted in Fig. 3(a); a total negative GVD of $-5,200\text{fs}^2$ per round-trip and a 7\% transmission output coupler.
Soliton mode-locked (ML) operation is observed for diode currents greater than 23A, corresponding to a continuous-wave (cw) output power of 1.8W.
Between diode currents (output powers) of 23 and 25A (1.8 and 2.3W), the resonator operates mainly in the cw mode, but will transition to ML operation upon being subjected to a perturbation.
The switching to ML operation is accompanied by a 250mW increase in output
%page 8637
power indicating a higher cavity Q for ML operation.
This is consistent both with the laser accessing more of the available inhomogeneous gain bandwidth and with our resonator design favoring operation with a strong Kerr lens in the Yb:KGW gain medium.
Given a value of $n^2 \approx 15 \cdot 10^{-16}\text{cm}^2/\text{W}$ for 1040nm radiation polarized parallel to the KGW crystal's optical $N_p$-axis\cite{Major_characterization_2003,Selivanov_nonlinear_2006,Vodchits_zscan_2006} and a calculated peak pulse irradiance of $\sim$4GW/cm$^2$ in the gain medium at 24A, we expect a Kerr lens strength of approximately 2m$^-1$.
From Fig. 3(b), we see that the resulting increase in the strength of the lensing in the Yb:KGW crystal will result in a reduction of the laser mode size in the gain medium, which produces a better spatial match to gain duct produced by the laser diode pumps (less absorptive loss), thus generating an increase in output power.

@@

The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.
At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
After the collimating
%page 8639
mirror, a standard flat dichroic beamsplitter then separates the reflected $s$-polarized 523nm radiation from the transmitted $p$-polarized fundamental.
